---
title: "UFC Fight History - ARINMA"
author: "Gregory Schlett"
date: "2025-04-04"
output: html_document
--- 

# UFC Fight History
## Data from: Kaggle
### Updated by: Alex Magnus
####### End

```{r}
getwd()
```

```{r}
# Uploading CSV file from working directory
ufc_data <- read.csv("UFC Fight Statistics (July 2016 - Nov 2024).csv")
# ufc_data <- read.csv("/Users/mreknoian/Desktop/UFC Fight Statistics (July 2016 - Nov 2024).csv")

# View of the first few rows of Fight Statistics
head(ufc_data)
```

```{r}
str(ufc_data)
```

```{r}
ufc_data$TotalStrikes.F1 <- ufc_data$Total.Strike.Landed.F1R1 + ufc_data$Total.Strike.Landed.F1R2 + ufc_data$Total.Strike.Landed.F1R3 + ufc_data$Total.Strike.Landed.F1R4 + ufc_data$Total.Strike.Landed.F1R5

ufc_data$TotalStrikes.F2 <- ufc_data$Total.Strike.Landed.F2R1 + ufc_data$Total.Strike.Landed.F2R2 + ufc_data$Total.Strike.Landed.F2R3 + ufc_data$Total.Strike.Landed.F2R4 + ufc_data$Total.Strike.Landed.F2R5

ufc_data$TotalSigStrikes.F1 <- ufc_data$Sig..Strike.Landed.F1R1 + ufc_data$Sig..Strike.Landed.F1R2 + ufc_data$Sig..Strike.Landed.F1R3 + ufc_data$Sig..Strike.Landed.F1R4 + ufc_data$Sig..Strike.Landed.F1R5
ufc_data$TotalSigStrikes.F2 <- ufc_data$Sig..Strike.Landed.F2R1 + ufc_data$Sig..Strike.Landed.F2R2 + ufc_data$Sig..Strike.Landed.F2R3 + ufc_data$Sig..Strike.Landed.F2R4 + ufc_data$Sig..Strike.Landed.F2R5

ufc_data$TotalKnockdowns.F1 <- ufc_data$Knockdowns.F1R1 + ufc_data$Knockdowns.F1R2 + ufc_data$Knockdowns.F1R3 + ufc_data$Knockdowns.F1R4 + ufc_data$Knockdowns.F1R5

ufc_data$TotalKnockdowns.F2 <- ufc_data$Knockdowns.F2R1 + ufc_data$Knockdowns.F2R2 + ufc_data$Knockdowns.F2R3 + ufc_data$Knockdowns.F2R4 + ufc_data$Knockdowns.F2R5

ufc_data$TotalTakedowns.F1 <- ufc_data$TD.Completed.F1R1 + ufc_data$TD.Completed.F1R2 + ufc_data$TD.Completed.F1R3 + ufc_data$TD.Completed.F1R4 + ufc_data$TD.Completed.F1R5

ufc_data$TotalTakedowns.F2 <- ufc_data$TD.Completed.F2R1 + ufc_data$TD.Completed.F2R2 + ufc_data$TD.Completed.F2R3 + ufc_data$TD.Completed.F2R4 + ufc_data$TD.Completed.F2R5

ufc_data$TotalControlTime.F1 <- ufc_data$Ctrl.Time..Minutes..F1R1 + ufc_data$Ctrl.Time..Minutes..F1R2 + ufc_data$Ctrl.Time..Minutes..F1R3 + ufc_data$Ctrl.Time..Minutes..F1R4 + ufc_data$Ctrl.Time..Minutes..F1R5

ufc_data$TotalControlTime.F2 <- ufc_data$Ctrl.Time..Minutes..F2R1 + ufc_data$Ctrl.Time..Minutes..F2R2 + ufc_data$Ctrl.Time..Minutes..F2R3 + ufc_data$Ctrl.Time..Minutes..F2R4 + ufc_data$Ctrl.Time..Minutes..F2R5
```

```{r}
#install.packages("forecast")
#install.packages("tseries")  # for testing stationarity
library(forecast)
library(tseries)
```

```{r}
ufc_data$Date <- as.Date(ufc_data$Date, format = "%m/%d/%Y")

ufc_data <- ufc_data[order(ufc_data$Date), ]

sig_ts <- ts(na.omit(ufc_data$TotalSigStrikes.F1), frequency = 1)

sig_ts
```

```{r}
adf.test(sig_ts)
```

```{r}
model <- auto.arima(sig_ts)
summary(model)
```

```{r}
future <- forecast(model, h = 10)
plot(future)
```

```{r}
checkresiduals(model)
```

```{r}
# Total takedowns for Fighter 1
takedowns_ts <- ts(ufc_data$TotalTakedowns.F1, frequency = 1)
takedown_model <- auto.arima(takedowns_ts)
forecast_takedown <- forecast(takedown_model, h = 10)
plot(forecast_takedown)
```

```{r}
accuracy(model)
```

# Summary of ARIMA
## ME - Unpredicts by 3.6 Sig Strikes
## RMSE - Large value indicates it sometimes makes large mistakes with a heavy penalty for outliers
## MAE - Averages an error of 35 significant strikes per fight (this is can be extremely influential in fights)
## MPE - Model is unpredicting outcomes
## MAPE - Off by more than half the actual value (Very High)
## MASE - Performs better than Naive
##ACF1 - Mild autocorrelation
### ARIMA model is better than a basic guess, but it has limited accuracy. It tends to underpredict, and it's often off by a large margin, especially in fights where the number of strikes is unusually high.

